---
title: "Frequency and character complexity in scripts"
author: "Alexey Koshevoy, Helena Miton and Olivier Morin"
date: "10/9/2021"
output: 
  html_document:
    theme: flatly
    number_sections: true
---

<style type="text/css">
  body{
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot) 
library(effects)
library(showtext)
library(ggeffects)
library(ggpubr)
library(lingtypology)
library(ggrepel)
library(MuMIn)
library(Cairo)
library(sysfonts)
library(broom.mixed)

theme_set(theme_bw())
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning = FALSE)
```

# Data loading

Dataframe:

```{r}
data <- read.csv('final.csv')
data <- data[data$ISO_language != 'heb',]

DT::datatable(data, 
              filter = 'top',
              rownames = FALSE,
              escape=FALSE, 
              # fontSize = '50%',
              extensions = 'FixedColumns',
  options = list(
  dom = 't',
  scrollX = TRUE,
  scrollCollapse = TRUE
)) %>%
  DT::formatStyle(columns = colnames(.),
                  fontSize = '20%')
```

Counting types of writing systems:

```{r}
data %>%
  distinct(Type, ISO_language)  %>%
  group_by(Type) %>% 
  tally()
```


Plot the map:

```{r}
dist <- data %>%
  distinct(ISO_script, .keep_all = TRUE)

dist$Language <- lang.iso(dist$ISO_language)
dist$lat <- lat.lang(dist$Language)
dist$lon <- long.lang(dist$Language)
dist[dist$ISO_language == 'chr',]$lat <- 35.8513
dist[dist$ISO_language == 'chr',]$lon <- -94.9878

dist %>%
  ggplot(aes(x=lon, y=lat))+
    coord_sf(xlim = c(-100, 120),ylim = c(0, 60), expand = TRUE)+
    borders("world", colour=alpha("gray50", .2), fill=alpha("gray50", .2))+
    geom_point(aes(color=Family))+
    geom_text_repel(aes(label=ISO_script))+
    theme_void()+
    # theme(legend.position = "none")+
    theme(legend.position = 'bottom')

ggsave('figures/fig1_map.png', dpi=300, width = 8, height = 6)

knitr::plot_crop('figures/fig1_map.png')
```

Count of symbols and number of elements in each WS:

```{r}
totals <- data %>%
  group_by(ISO_script, ISO_language) %>%
  summarise(sum=sum(Frequency), n=n()) %>%
  arrange(desc(n)) 

DT::datatable(totals, 
              filter = 'top',
              rownames = FALSE,
              escape=FALSE, 
              extensions = 'FixedColumns',
  options = list(
  dom = 't',
  scrollX = TRUE,
  scrollCollapse = TRUE
))
```

Median values for n of characters in the corpus and n of symbols:

```{r}
median(totals$n)
median(totals$sum)
```

Log-transformation:

```{r}
nlog <- ggplot(data, aes(x=Relative_frequency))+
  theme(legend.position = "none")+
  geom_histogram(aes(y = ..density..), bins=50)+
  geom_density(color='red', lwd=1)+
  # ggtitle('Initial frequencies')+
  xlab('Relative frequency')

log <- ggplot(data, aes(x=log(Relative_frequency)))+
  theme(legend.position = "none")+
  geom_histogram(aes(y = ..density..), bins=50)+
  geom_density(color='red', lwd=1)+
  # ggtitle('Log-transformed frequencies')+
  xlab('Relative frequency (log-transformed)')

ggarrange(nlog, log)
```

Individual distributions:

```{r}
ggplot(data, aes(x=log(Relative_frequency), color=ISO_language))+
  theme(legend.position = "none")+
  geom_histogram()+
  facet_wrap(~ISO_language, scale='free_y')
```

Correlation of complexity measures:

```{r}
ggplot(data, aes(x=Perimetric_complexity, y=Compression))+
  theme(legend.position = "none")+
  geom_point(size=0.5)+
  facet_wrap(~ISO_language)
```

# Analysis

Log-transforming relative frequency:

```{r}
data$Relative_frequency_l <- log(data$Relative_frequency)
```


Defining the model and comparing the AIC values:

```{r}
# Perimetric complexity model:
model_full_p <- lmer(formula = Perimetric_complexity ~ 1 + Relative_frequency_l + (1 + Relative_frequency_l|ISO_script), data=data)
# Algorithmic complexity model:
model_full_c <- lmer(formula = Compression ~ 1 + Relative_frequency_l + (1 + Relative_frequency_l|ISO_script), data=data)
model_null_p <- lmer(formula =Perimetric_complexity ~ 1 + (1|ISO_script), data=data)
model_null_c <- lmer(formula =Compression ~ 1 + (1|ISO_script), data=data)
AIC(model_null_p) - AIC(model_full_p)
AIC(model_null_c) - AIC(model_full_c)
```

Model with Perimetric Complexity:

```{r}
summary(model_full_p)
```

```{r}
p_sim <- data.frame(Relative_frequency_l = seq(min(data$Relative_frequency_l), max(data$Relative_frequency_l), .01)
                     )
fit.mat <- model.matrix(~ Relative_frequency_l, p_sim) 
cis <- diag(fit.mat %*% tcrossprod(vcov(model_full_p), fit.mat))

p_sim$Perimetric_complexity <- predict(model_full_p, p_sim, re.form = NA)
p_sim$lwr <- p_sim$Perimetric_complexity-1.96*sqrt(cis)
p_sim$upr <- p_sim$Perimetric_complexity+1.96*sqrt(cis)
```

```{r}
script <- 'Thai'
font_add("thai", "fonts/thai.ttf")

p1 <- data %>%
  filter(ISO_script == script) %>%
  ggplot(aes(x=Relative_frequency_l, y=Perimetric_complexity))+
  geom_point(data=data,
             aes(x=Relative_frequency_l,
                 y=Perimetric_complexity),
             alpha=0.08)+
  geom_ribbon(data=p_sim,
              aes(x=Relative_frequency_l,
                  y=Perimetric_complexity,
                  ymin = lwr,
                  ymax = upr),
              alpha = .3,
              fill = "red") +
  geom_line(data=p_sim,
            aes(x=Relative_frequency_l,
                y=Perimetric_complexity),
            size = 1, color = "red")+
  geom_point()+
  geom_text_repel(
    aes(label=stringi::stri_enc_toutf8(textfile)),
    family = 'thai',
    nudge_x = 0.25, nudge_y = 0.25
  )+
  scale_y_continuous(breaks = seq(0, max(data$Perimetric_complexity), by = 20))+
  scale_x_continuous(breaks = seq(-15, 0, by = 3))+
  xlab('Relative frequency (log-transformed)')+
  ylab('Perimetric complexity')
```


Model with compression:

```{r}
summary(model_full_c)
```


```{r}
p_sim <- data.frame(Relative_frequency_l = seq(min(data$Relative_frequency_l), max(data$Relative_frequency_l), .01)
                     )
fit.mat <- model.matrix(~ Relative_frequency_l, p_sim) 
cis <- diag(fit.mat %*% tcrossprod(vcov(model_full_c), fit.mat))

p_sim$Compression <- predict(model_full_c, p_sim, re.form = NA)
p_sim$lwr <- p_sim$Compression-1.96*sqrt(cis)
p_sim$upr <- p_sim$Compression+1.96*sqrt(cis)

script <- 'Mlym'
font_add("mlym", "fonts/malaylam.ttf")

p2 <- data %>%
  filter(ISO_script == script) %>%
  ggplot(aes(x=Relative_frequency_l, y=Compression))+
  geom_point(data=data,
             aes(x=Relative_frequency_l,
                 y=Compression),
             alpha=0.08)+
  geom_ribbon(data=p_sim,
              aes(x=Relative_frequency_l,
                  y=Compression,
                  ymin = lwr,
                  ymax = upr),
              alpha = .3,
              fill = "blue") +
  geom_line(data=p_sim,
            aes(x=Relative_frequency_l,
                y=Compression),
            size = 1, color = "blue")+
  geom_point()+
  geom_text_repel(
    aes(label=stringi::stri_enc_toutf8(textfile)),
    family = 'mlym',
    nudge_x = 0.25, nudge_y = 0.25
  )+
  scale_y_continuous(breaks = seq(0, max(data$Compression), by = 200))+
  scale_x_continuous(breaks = seq(-15, 0, by = 3))+
  xlab('Relative frequency (log-transformed)')+
  ylab('Algorithmic complexity')
```

```{r}
ggarrange(p1, p2, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

ggsave('figures/fig2_pred.png',
       width = 8,
       height = 4)
```



R-squared PC:

```{r}
r.squaredGLMM(model_full_p)
```

R-squarred AC:

```{r}
r.squaredGLMM(model_full_c)
```

Confidence intervals:

```{r}
confint(model_full_p)
```

```{r}
confint(model_full_c)
```

Get predicted values:

```{r}
predicted_values_c <- modelr::data_grid(data, Relative_frequency_l, ISO_script) %>% 
                   modelr::add_predictions(model_full_c)

predicted_values_p <- modelr::data_grid(data, Relative_frequency_l, ISO_script) %>% 
                   modelr::add_predictions(model_full_p)

predicted_values_c
```

Plot residuals:


```{r}
ggplot(data.frame(x=log(data$Relative_frequency),pearson=residuals(model_full_p,type="pearson")),
    aes(x=x,y=pearson)) +
    geom_point() +
    theme_bw()+
    xlab('Relative frequency')
```


```{r}
qqnorm(residuals(model_full_p))
```


```{r}
qqnorm(residuals(model_full_c))
```


```{r}
ggplot(augment(model_full_c), aes(.fitted, .resid))+
  geom_point(alpha=0.5)+
  geom_hline(yintercept=0, col="red", linetype="dashed")+
  xlab("Fitted values")+
  ylab("Residuals")
```

```{r}
ggplot(augment(model_full_p), aes(.fitted, .resid))+geom_point(alpha=0.5)+
  geom_hline(yintercept=0, col="red", linetype="dashed")+
  xlab("Fitted values")+
  ylab("Residuals")+
  ylim(c(-60, 60))
```


Perimetric complexity predicitons:

```{r}
predicted_values_p %>% 
  ggplot(aes(Relative_frequency_l, pred, color = ISO_script))+
  geom_point(data=data, aes(x=Relative_frequency_l, y=Perimetric_complexity, color=ISO_script), alpha=0.2)+
  geom_line()+
  theme(legend.position = "none")+
  xlab('Frequency (log-transformed)')+
  ylab('Perimetric complexity')+
  facet_wrap(~ISO_script, scale='free_x')

ggsave('figures/fig1_p.png')
```

Compression predictions:

```{r}
predicted_values_c %>% 
  ggplot(aes(Relative_frequency_l, pred, color = ISO_script))+
  geom_point(data=data, aes(x=Relative_frequency_l, y=Compression, color=ISO_script), alpha=0.2)+
  geom_line()+
  theme(legend.position = "none")+
  xlab('Frequency (log-transformed)')+
  ylab('Algorithmic complexity')+
  facet_wrap(~ISO_script, scale='free_x')

ggsave('figures/fig1_c.png')
```

Predictions perimetric complexity:

```{r}
predicted_values_p %>% 
  ggplot(aes(Relative_frequency_l, pred, color = ISO_script))+
  geom_point(data=data, aes(x=Relative_frequency_l, y=Perimetric_complexity, color=ISO_script), alpha=0.2, show.legend = F)+
  geom_line()+
  theme(legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_blank())+
  xlab('Frequency (log-transformed)')+
  ylab('Perimetric complexity')+
  guides(col=guide_legend(nrow=3, byrow=TRUE))

ggsave('figures/fig2_p.png',  width = 12, height = 7)
```


Predictions compression:

```{r}
predicted_values_c %>% 
  ggplot(aes(Relative_frequency_l, pred, color = ISO_script))+
  geom_point(data=data, aes(x=Relative_frequency_l, y=Compression, color=ISO_script), alpha=0.2, show.legend = F)+
  geom_line()+
  theme(legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_blank())+
  xlab('Frequency (log-transformed)')+
  ylab('Compression')+
  guides(col=guide_legend(nrow=3, byrow=TRUE))

ggsave('figures/fig2_c.png')
```
# Miscellaneous

```{r}
library(merTools)
randomSims <- REsim(model_full_c, n.sims = 10000)
randomSims[randomSims$term == 'Relative_frequency_l',]  %>%
  full_join(totals, by=c('groupID' = 'ISO_script')) %>%
  mutate(groupID = fct_reorder(groupID, desc(mean))) %>%
  mutate(error=qnorm(0.975)*sd/sqrt(n)) %>%
  mutate(left=mean-error, right=mean+error)%>%
  ggplot(aes(y=groupID, x=mean))+
  geom_point()+
  geom_errorbar(aes(xmin=left, xmax=right))+
  geom_vline(xintercept = 0, color='red', linetype='dashed')

# randomSims
```


```{r}
randomSims <- REsim(model_full_p, n.sims = 10000)
randomSims[randomSims$term == 'Relative_frequency_l',]  %>%
  full_join(totals, by=c('groupID' = 'ISO_script')) %>%
  mutate(groupID = fct_reorder(groupID, desc(mean))) %>%
  mutate(error=qnorm(0.975)*sd/sqrt(n)) %>%
  mutate(left=mean-error, right=mean+error)%>%
  ggplot(aes(y=groupID, x=mean))+
  geom_point()+
  geom_errorbar(aes(xmin=left, xmax=right))+
  geom_vline(xintercept = 0, color='red', linetype='dashed')
```


```{r}
randomSims[randomSims$term == 'Relative_frequency_l',] %>%
  full_join(totals, by=c('groupID' = 'ISO_script')) %>%
  ggplot(aes(x=n, y=mean))+
  geom_point()
```


# Individual predictions

```{r}
# font_add_google('Nanum Gothic', 'nanum')
# font_add_google('Benne', 'ben')
# font_add_google('Noto', 'noto', db_cache = FALSE)
# font_add_google('Noto Serif Georgian', 'georg')
font_add("georgian", "fonts/georg.ttf")

showtext_auto()

script <- 'Geor'

data %>%
  filter(ISO_script == script) %>%
  ggplot(aes(x=Relative_frequency_l, y=Perimetric_complexity))+
  geom_line(data=predicted_values_p[predicted_values_p$ISO_script == script,], aes(x=Relative_frequency_l, y=pred))+
  geom_text(
    aes(label=stringi::stri_enc_toutf8(textfile)),
    family = 'georgian',
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )

ggsave('try.png', width = 6, height = 3)
```

```{r}
script <- 'Thai'
font_add("thai", "fonts/thai.ttf")

data %>%
  filter(ISO_script == script) %>%
  ggplot(aes(x=Relative_frequency_l, y=Perimetric_complexity))+
  geom_line(data=predicted_values_p[predicted_values_p$ISO_script == script,], aes(x=Relative_frequency_l, y=pred))+
  # theme(text=element_text(family="nanum"))+
  geom_text(
    aes(label=stringi::stri_enc_toutf8(textfile)),
    family = 'thai',
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )+
  xlab('Frequency (log-transformed)')+
  ylab('Perimetric complexity')

ggsave('figures/example1.png', width = 6, height = 3)
```

```{r}
script <- 'Mlym'
font_add("mlym", "fonts/malaylam.ttf")

data %>%
  filter(ISO_script == script) %>%
  ggplot(aes(x=Relative_frequency_l, y=Compression))+
  geom_line(data=predicted_values_c[predicted_values_p$ISO_script == script,], aes(x=Relative_frequency_l, y=pred))+
  # theme(text=element_text(family="nanum"))+
  geom_text(
    aes(label=stringi::stri_enc_toutf8(textfile)),
    family = 'mlym',
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )+
  xlab('Frequency (log-transformed)')+
  ylab('Algorithmic complexity')

ggsave('figures/example2.png', width = 6, height = 3)
```

```{r}
script <- 'Thai'
font_add("thai", "fonts/thai.ttf")

data %>%
  filter(ISO_script == script) %>%
  ggplot(aes(x=Relative_frequency_l, y=Compression))+
  geom_line(data=predicted_values_c[predicted_values_p$ISO_script == script,], aes(x=Relative_frequency_l, y=pred))+
  # theme(text=element_text(family="nanum"))+
  geom_text(
    aes(label=stringi::stri_enc_toutf8(textfile)),
    family = 'thai',
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )+
  xlab('Frequency (log-transformed)')+
  ylab('Algorithmic complexity')

ggsave('figures/example3.png', width = 6, height = 3)
```



